---
title: "220822_YoderSC_06_celltypes"
author: "T.C. Howton"
date: '2022-08-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=10, fig.height=20) 
knitr::opts_chunk$set(fig.width=15, fig.height=10) 
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
    library(Seurat)
    library(dplyr)
    library(cowplot)
    library(ggplot2)
    library(pheatmap)
    library(enrichR)
    library(rafalib)
})
```


### Load data

```{r Load-Data}
# Loading in the filtered seurat object from the dge script "/data/user/tchowton/220804_YoderSC/Scripts/220822_YoderSC_05_dge.Rmd"
alldata <- readRDS("/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220822_YoderSC_05-ift88-dge.rds")

dim(alldata) #20405 20625
print(names(alldata@reductions)) # "pca"          "umap"         "harmony"      "umap_harmony"
```

### Initial Marker Annotation

```{r}
# https://www.science.org/doi/10.1126/science.aar2131
cd_genes <- c('Nrp1', 'Kdr', 'Nphs1', 'Nphs2', 'Slc27a2', 'Lrp2', 'Slc12a1', 'Umod', 'Slc12a3', 'Pvalb', 'Aqp2', 'Hsd11b2', 'Atp6v1g3', 'Atp6v0d2', 'Insrr', 'Rhbg', 'Mki67', 'Cdca3', 'Plac8', 'S100a4', 'C1qa', 'C1qb', 'S100a8', 'S100a9', 'Cd79a', 'Cd79b', 'Ltb', 'Cxcr6', 'Gzma', 'Nkg7', 'Stmn1')
pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220823_res-5-celltype-markers.pdf', width = 28, height = 18)
DotPlot(object = alldata, features = cd_genes)
dev.off()

# Cell counts by cluster
#   1    2    3    4    5    6    7    8    9   10   11   12   13   14   15   16   17   18   19   20   21   22   23   24   25   26   27   28 
#2916 2185 1727 1396 1344 1300 1188  977  923  922  876  692  597  425  377  338  335  308  307  299  290  262  234  132   91   68   60   56 
```

```{r Vlnplot-per-marker}
pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220822_celltype-markers-individual.pdf')
for (i in 1:length(cd_genes)) {
  print(VlnPlot(alldata, features = cd_genes[[i]], pt.size = 0))
}
dev.off()
```


### Panglaodb Check for Unknown Clusters
Some clusters appeared to be a certain cell type from the initial check, but I want to look at the top few genes to double check. 

```{r Individual-cluster-id}
# Cluster 5 didn't have any markers pop up so using https://panglaodb.se/ to check the top genes
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 5], F),
        horiz = T, las = 1, main = paste0(5, " vs. rest"), border = "white", yaxs = "i")
# cluster 7
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 7], F),
        horiz = T, las = 1, main = paste0(7, " vs. rest"), border = "white", yaxs = "i")
# T cells

# cluster 9
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 9], F),
        horiz = T, las = 1, main = paste0(9, " vs. rest"), border = "white", yaxs = "i")
# Macrophages?

# cluster 10
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 10], F),
        horiz = T, las = 1, main = paste0(10, " vs. rest"), border = "white", yaxs = "i")
# Macrophages?

# cluster 19
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 19], F),
        horiz = T, las = 1, main = paste0(19, " vs. rest"), border = "white", yaxs = "i")
# T cells

# cluster 22
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 22], F),
        horiz = T, las = 1, main = paste0(22, " vs. rest"), border = "white", yaxs = "i")
# Macrophages or Fibroblasts

# cluster 31
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 31], F),
        horiz = T, las = 1, main = paste0(31, " vs. rest"), border = "white", yaxs = "i")
# Macrophages

# cluster 37
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 37], F),
        horiz = T, las = 1, main = paste0(37, " vs. rest"), border = "white", yaxs = "i")
# PTC

# cluster 41
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 41], F),
        horiz = T, las = 1, main = paste0(41, " vs. rest"), border = "white", yaxs = "i")
# B cells (still weird its so far away)

# cluster 48
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 48], F),
        horiz = T, las = 1, main = paste0(48, " vs. rest"), border = "white", yaxs = "i")
# Endothelial cells

# cluster 49
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 49], F),
        horiz = T, las = 1, main = paste0(49, " vs. rest"), border = "white", yaxs = "i")
# macrophages (still check)

# cluster 52
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 52], F),
        horiz = T, las = 1, main = paste0(52, " vs. rest"), border = "white", yaxs = "i")
# PTC

# cluster 54
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 54], F),
        horiz = T, las = 1, main = paste0(54, " vs. rest"), border = "white", yaxs = "i")
# T cells

# cluster 56
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 56], F),
        horiz = T, las = 1, main = paste0(56, " vs. rest"), border = "white", yaxs = "i")
# Endothelial cells

# cluster 50
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 50], F),
        horiz = T, las = 1, main = paste0(50, " vs. rest"), border = "white", yaxs = "i")
# Fibroblasts (smc)

# cluster 39
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 39], F),
        horiz = T, las = 1, main = paste0(39, " vs. rest"), border = "white", yaxs = "i")
# DTC?

# cluster 43
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 43], F),
        horiz = T, las = 1, main = paste0(43, " vs. rest"), border = "white", yaxs = "i")
# macrophages (still check)

# cluster 51
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 51], F),
        horiz = T, las = 1, main = paste0(51, " vs. rest"), border = "white", yaxs = "i")
# Fib (still cehck)

# cluster 53
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 53], F),
        horiz = T, las = 1, main = paste0(53, " vs. rest"), border = "white", yaxs = "i")
# smooth muscle cells

# cluster 57
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 57], F),
        horiz = T, las = 1, main = paste0(57, " vs. rest"), border = "white", yaxs = "i")
# unknown

# cluster 58
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 58], F),
        horiz = T, las = 1, main = paste0(58, " vs. rest"), border = "white", yaxs = "i")
# unknown 2

# cluster 62
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 62], F),
        horiz = T, las = 1, main = paste0(62, " vs. rest"), border = "white", yaxs = "i")
#  unknown 3

# cluster 63
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 63], F),
        horiz = T, las = 1, main = paste0(63, " vs. rest"), border = "white", yaxs = "i")
# macrophages

# cluster 64
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 64], F),
        horiz = T, las = 1, main = paste0(64, " vs. rest"), border = "white", yaxs = "i")
# macrophages

# cluster 65
barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == 65], F),
        horiz = T, las = 1, main = paste0(65, " vs. rest"), border = "white", yaxs = "i")
# EC
```

### Cell Type Specific Checks

```{r cell-type-specific-checks}
DimPlot(alldata, label = T) + NoAxes()
# Macrophages
VlnPlot(alldata, features = "Gzma", pt.size = 0) 
VlnPlot(alldata, features = "Nkg7", pt.size = 0)
VlnPlot(alldata, features = "Adgre1", pt.size = 0) 
VlnPlot(alldata, features = "Fcgr1", pt.size = 0)
FeaturePlot(alldata, features = c("Itgam"))
FeaturePlot(alldata, features = c("Ccr2"))
FeaturePlot(alldata, features = c("Adgre1"))
FeaturePlot(alldata, features = c("Fcgr1")) #cd64

# Dendritic cells
VlnPlot(alldata, features = "Snx22", pt.size = 0) #inf 6, 13, 20
VlnPlot(alldata, features = "Cd209a", pt.size = 0)
FeaturePlot(alldata, features = c("Snx22"))
FeaturePlot(alldata, features = c("Batf3"))

# podocytes
VlnPlot(alldata_scaled, features = "Cd4", pt.size = 0) 
VlnPlot(alldata, features = "Cd4", pt.size = 0)
VlnPlot(alldata, features = "Cd3d", pt.size = 0)
VlnPlot(alldata, features = "Cd3g", pt.size = 0) 
VlnPlot(alldata, features = "Slc22a6", pt.size = 0)
VlnPlot(alldata, features = "Slc22a8", pt.size = 0)
VlnPlot(alldata, features = "Car4", pt.size = 0) 
VlnPlot(alldata, features = "Agt", pt.size = 0)
VlnPlot(alldata, features = "Lyz2", pt.size = 0)
VlnPlot(alldata, features = "Slc27a2", pt.size = 0) 
RidgePlot(alldata_subset, features = c("Nphs1", "Nphs2"))
RidgePlot(alldata_scaled, features = "Nphs1")
VlnPlot(alldata, features = "Cd14", pt.size = 0)
FeaturePlot(alldata, features = c("Snx22"))
FeaturePlot(alldata, features = c("Batf3"))

Kasi_markers <- c('Nphs1', 'Nphs2', 'Adamts5', 'Plat', 'Ehd3', 'Mapt', 'Plvap', 'Pdgfrb', 'Slc34a1', 'Lrp2', 'Slc5a2', 'Slc5a12', 'Slc22a6', 'Slc22a8', 'Car4', 'Agt', 'Slc5a1', 'Slc27a2', 'Slc22a19', 'Slc22a7', 'Slc12a1', 'Clcnka', 'Umod', 'Slc12a3', 'Aqp2', 'Scnn1g', 'Hsd11b2', 'Foxi1', 'Slc4a1', 'Aqp6', 'Slc26a4', 'Hmx2', 'Ms4a1', 'Cd79a', 'Jchain', 'Mzb1', 'Cd209a', 'Gzma', 'S100a9', 'Cd68', 'Adgre1', 'Itgam', 'C1qa', 'C1qb', 'C1qc', 'Apoe', 'Ly6c1', 'Ly6c2', 'Plac8', 'Chil3', 'Ear2', 'Cebpb', 'Nr4a1', 'Cd3d', 'Cd3g', 'Cd4', 'Cd8a')

DoHeatmap(alldata_scaled, features = Kasi_markers)

tubule_markers <- c('Slc5a2', 'Slc22a6', 'Lrp2', 'Slc9a3', 'Pth1r', 'Slc34a1', 'Agt', 'Aqp1', 'Slc14a2', 'Clcnka', 'Slc12a1', 'Umod', 'Pvalb', 'Slc12a3', 'Calb1', 'Aqp2', 'Scnn1g', 'Slc4a1', 'Slc26a4', 'Atp6c1g3')

DoHeatmap(alldata_scaled, features = tubule_markers)

DotPlot(object = alldata_scaled, features = tubule_markers)
DotPlot(object = alldata, features = tubule_markers)
for (i in Kasi_markers) {
  print(RidgePlot(alldata_subset, features = i))
}

alldata_scaled <- ScaleData(alldata)

DotPlot(object = alldata, features = Kasi_markers)

DotPlot(object = alldata, features = c("Pdgfra", "Gata3"))

macros <- c('Cd68', 'Adgre1', 'Itgam', 'C1qa', 'C1qb', 'C1qc', 'Apoe')
DotPlot(object = alldata_subset, features = macros)

alldata_subset <- subset(alldata, idents = c(57, 58))
DimPlot(alldata_subset, label = T) + NoAxes()

VlnPlot(alldata, features = "Pdgfrb", pt.size = 0)
FeaturePlot(alldata, features = c("Adamts5"))

```


### Merging and renaming clusters

```{r Merge-Rename-Clusters}
new_cluster_ids <- c('Cytotoxic_T_Cell', 'Helper_T_Cell', 'Endothelial_Cells', 'Resident_Macrophages', 'Endothelial_Cells', 'Resident_Macrophages', 'Helper_T_Cell', 'Endothelial_Cells', 'Infiltrating_Macrophages', 'Infiltrating_Macrophages', 'B_Lymphocytes', 'Resident_Macrophages', 'Natural Kill Cells', 'Endothelial_Cells', 'B_Lymphocytes', 'Endothelial_Cells', 'Endothelial_Cells', 'B_Lymphocytes', 'Helper_T_Cell', 'Endothelial_Cells', 'Resident_Macrophages', 'Infiltrating_Macrophages', 'Collecting_Duct_Principal_Cell', 'Proximal_Tubule_S2_Cells', 'Helper_T_Cell', 'Proximal_Tubule_S3_Cells', 'Distal_Convoluted_Tubule_Cells', 'Proximal_Tubule_S1_Cells', 'Resident_Macrophages', 'Loop_of_Henle', 'Infiltrating_Macrophages', 'Proximal_Tubule_S2_Cells', 'Helper_T_Cell', 'Endothelial_Cells', 'Resident_Macrophages', 'Resident_Macrophages', 'Proximal_Tubule_S2_Cells', 'Neutrophils', 'Collecting_Duct_Principal_Cell', 'Proximal_Tubule_S1_Cells', 'B_Lymphocytes', 'Resident_Macrophages', 'Dendritic_Cells', 'Cytotoxic_T_Cell', 'Proximal_Tubule_S1_Cells', 'Proximal_Tubule_S1_Cells', 'Collecting_Duct_Principal_Cell', 'Endothelial_Cells', 'Resident_Macrophages', 'Pericytes', 'Fibroblasts', 'Proximal_Tubule_S3_Cells', 'Smooth_Muscle_Cells', 'T_Lymphocytes', 'Helper_T_Cell', 'Resident_Macrophages', 'Unknown_1', 'Unknown_2', 'Collecting_Duct_Intercalated_Cell', 'Endothelial_Cells', 'Podocytes', 'Infiltrating_Macrophages', 'Resident_Macrophages', 'Resident_Macrophages', 'Endothelial_Cells')
names(new_cluster_ids) <- levels(alldata)
alldata <- RenameIdents(alldata, new_cluster_ids)
pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220825_labeled.pdf', width = 10, height = 10)
DimPlot(alldata, reduction = "umap_harmony", label = TRUE, pt.size = 0.5) + NoLegend()
dev.off
#alldata_distal <- subset(alldata, idents = "Distal_Tuble_Cells")
```


### Save Data
```{r SaveData-SessionInfo}
saveRDS(alldata, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220825_YoderSC_06-ift88-final.rds")

sessionInfo()
```