---
title: "220818_YoderSC_04_clustering"
author: "T.C. Howton"
date: '2022-08-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
library(Seurat)
library(cowplot)
library(ggplot2)
library(pheatmap)
library(rafalib)
library(clustree)
library(reticulate)
})
```


### Load data

```{r Load-Data}
# Loading in the filtered seurat object from the integration script "/data/user/tchowton/220804_YoderSC/Scripts/220817_YoderSC_03_integration.Rmd"
alldata <- readRDS("/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220817_YoderSC_03-ift88-integration.rds")

dim(alldata) #20405 20625
print(names(alldata@reductions)) # "pca"  "umap"
```

### Building kNN / SNN graph

The first step into graph clustering is to construct a k-nn graph.

```{r}
# check that CCA is still the active assay
alldata@active.assay <- "RNA"
alldata@active.assay

alldata <- FindNeighbors(alldata, dims = 1:30, reduction = "harmony")

## Computing nearest neighbor graph

## Computing SNN

# check the names for graphs in the object.
names(alldata@graphs) # "RNN_nn" "RNA_snn

#pheatmap(alldata@graphs$RNA_nn[1:200, 1:200],col = c("white", "black"), border_color = "grey90",legend = F, cluster_rows = F, cluster_cols = F, fontsize = 2)

```

### Clustering with Leiden Algorithm

```{r}
# Clustering with leiden (algorithm 4)
for (res in c(0.1, 0.25, 0.5, 1, 1.5, 2, 5)) {
    alldata <- FindClusters(alldata, graph.name = "RNA_snn", resolution = res, algorithm = 4, method = "igraph")
}

pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220822_clust-by-res.pdf')
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.0.5")
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.1")
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.2")
dev.off()

pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220823_clust-by-res.pdf')
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.0.5")
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.1")
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.2")
DimPlot(alldata, reduction = "umap", group.by = "RNA_snn_res.5")
dev.off()
```

### Clustree
We can now use the clustree package to visualize how cells are distributed between clusters depending on resolution.

```{r Clustree}
pdf('/data/user/tchowton/220804_YoderSC/Figures/ift88-clustering/220823_clustree.pdf')
clustree(alldata@meta.data, prefix = "RNA_snn_res.")
dev.off()
```

### Save Data
```{r SaveData-SessionInfo}
saveRDS(alldata, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220822_YoderSC_04-ift88-clustering.rds")

sessionInfo()
```






















