---
title: "220817_YoderSC_03_integration"
author: "T.C. Howton"
date: '2022-08-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(cowplot)
  library(ggplot2)
  library(harmony)
})

```


### Load data

```{r Load-Data}
# Loading in the filtered seurat object from the dim reduction script "/data/user/tchowton/220804_YoderSC/Scripts/220816_YoderSC_02_dim-reduction.Rmd"
alldata <- readRDS("/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220829_YoderSC_02-pc2-dr.rds")

dim(alldata) #20405 20625
print(names(alldata@reductions)) # "pca"  "umap"

# pc2
# [1] 19893 10688
# [1] "pca"  "umap"
```

### Split Seurat object into each sample and append to a list
We split the combined object into a list, with each dataset as an element. 

```{r Split-Seurat-Obj}
alldata.list <- SplitObject(alldata, split.by = "orig.ident")
```

### Normalize and Scale Each Obj
We perform standard preprocessing (log-normalization), and identify variable features individually for each dataset based on a variance stabilizing transformation (vst).

```{r Normalize-Scale}
for (i in 1:length(alldata.list)) {
    alldata.list[[i]] <- NormalizeData(alldata.list[[i]], verbose = FALSE)
    alldata.list[[i]] <- FindVariableFeatures(alldata.list[[i]], selection.method = "vst",
        nfeatures = 2000, verbose = FALSE)
}

hvgs_per_dataset <- lapply(alldata.list, function(x) {
    x@assays$RNA@var.features
})
# venn::venn(hvgs_per_dataset,opacity = .4,zcolor = scales::hue_pal()(3),cexsn
# = 1,cexil = 1,lwd=1,col='white',frame=F,borders = NA)

temp <- unique(unlist(hvgs_per_dataset))
overlap <- sapply(hvgs_per_dataset, function(x) {
    temp %in% x
})
pheatmap::pheatmap(t(overlap * 1), cluster_rows = F, color = c("grey90", "grey20"))
```

### Find-Int-Anchors
We identify anchors using the FindIntegrationAnchors function, which takes a list of Seurat objects as input.

```{r Find-Int-Anchors}
alldata.anchors <- FindIntegrationAnchors(object.list = alldata.list,
                                          dims = 1:30,
                                          reduction = "cca")
```

### Data-Integration-CCA
We then pass these anchors to the IntegrateData function, which returns a Seurat object.

```{r Int-Data-CCA}
alldata.int <- IntegrateData(anchorset = alldata.anchors, dims = 1:30, new.assay.name = "CCA")

names(alldata.int@assays) # "RNA" "CCA"

# by default, Seurat now sets the integrated assay as the default assay, so any
# operation you now perform will be on the ingegrated data.

alldata.int@active.assay # "CCA"
```

### Run Dimensionality reduction on integrated space CCA

```{r DR-CCA}
alldata.int <- ScaleData(alldata.int, verbose = FALSE)
alldata.int <- RunPCA(alldata.int, npcs = 30, verbose = FALSE)
alldata.int <- RunUMAP(alldata.int, dims = 1:30)
```

```{r DR-CCA-plot}
plot_grid(ncol = 2,
  DimPlot(alldata, reduction = "pca", group.by = "orig.ident")+NoAxes()+ggtitle("PCA raw_data"),
  DimPlot(alldata, reduction = "umap", group.by = "orig.ident")+NoAxes()+ggtitle("UMAP raw_data"),
  DimPlot(alldata.int, reduction = "pca", group.by = "orig.ident")+NoAxes()+ggtitle("PCA CCA integrated"),
  DimPlot(alldata.int, reduction = "umap", group.by = "orig.ident")+NoAxes()+ggtitle("UMAP CCA integrated")
)
```

### Harmony-Int
```{r}
alldata.harmony <- RunHarmony(alldata,
                              group.by.vars = "orig.ident",
                              reduction = "pca",
                              dims.use = 1:50, assay.use = "RNA")

# Here we use all PCs computed from Harmony for UMAP calculation

alldata.int[["harmony"]] <- alldata.harmony[["harmony"]]
alldata.int <- RunUMAP(alldata.int, dims = 1:50, reduction = "harmony", reduction.name = "umap_harmony")
```

### Integration Plotting for CCA and Harmony
```{r}
p1 <- DimPlot(alldata, reduction = "umap", group.by = "orig.ident") + ggtitle("UMAP raw_data")
p2 <- DimPlot(alldata.int, reduction = "umap", group.by = "orig.ident") + ggtitle("UMAP CCA")
p3 <- DimPlot(alldata.int, reduction = "umap_harmony", group.by = "orig.ident") +
    ggtitle("UMAP Harmony")
leg <- get_legend(p1)

pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-preprocessing/220829_int-cca-harmony-umap.pdf')
gridExtra::grid.arrange(gridExtra::arrangeGrob(p1 + NoLegend() + NoAxes(), p2 + NoLegend() +
    NoAxes(), p3 + NoLegend() + NoAxes(), nrow = 2),
    leg, ncol = 2, widths = c(8, 2))
dev.off()

DimPlot(alldata.int,
        reduction = "umap_harmony",
        split.by = "orig.ident") +
    ggplot2::ggtitle(label = "UMAP Harmony")

```

### Save Data
```{r SaveData-SessionInfo}
saveRDS(alldata.int, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220829_YoderSC_03-pc2-integration.rds")

sessionInfo()
```







