---
title: "220825_YoderSC_07_liana"
author: "T.C. Howton"
date: '2022-08-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=15, fig.height=10) #nichenet-1
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
    library(tidyverse)
    library(nichenetr)
    library(Seurat)
    library(ggrepel)
    library(cowplot)
})

# System has not been booted with systemd as init system (PID 1). Can't operate.
# Failed to create bus connection: Host is down
# Warning in system("timedatectl", intern = TRUE) :
  # running command 'timedatectl' had status 1
```


### Load data

```{r Load-Data}
# Loading in the final seurat object from the celltypes script "/data/user/tchowton/220804_YoderSC/Scripts/220822_YoderSC_06_celltypes.Rmd"
alldata <- readRDS("/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220825_YoderSC_06-ift88-final.rds")

dim(alldata) #20405 20625

# Set cell types
alldata$celltype <- alldata@active.ident

# Double check meta.data
alldata@meta.data %>% head()

DimPlot(alldata, label = T) + NoAxes() +NoLegend()

```

### Read in NicheNet Priors

```{r}
# The URL method did not work so I downloaded from the web and uploaded into Data
ligand_target_matrix <- readRDS("/data/user/tchowton/220804_YoderSC/Data/NicheNetPrior/ligand_target_matrix.rds")
lr_network <- readRDS("/data/user/tchowton/220804_YoderSC/Data/NicheNetPrior/lr_network.rds")
weighted_networks <- readRDS("/data/user/tchowton/220804_YoderSC/Data/NicheNetPrior/weighted_networks.rds")

weighted_networks_lr <- weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))

lr_network <- lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) <- ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix) <- ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()

ligand_target_matrix <- ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]

weighted_networks_lr <- weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```


## Run NicheNet

### Assign sender and receiver clusters

```{r}
alldata@meta.data$celltype
## receiver
receiver <- "Resident_Macrophages"
expressed_genes_receiver <- get_expressed_genes(receiver, alldata, pct = 0.10)

background_expressed_genes <- expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]

## sender
sender_celltypes <- c("Proximal_Tubule_S1_Cells", "Proximal_Tubule_S1_Cells", "Proximal_Tubule_S3_Cells")

list_expressed_genes_sender <- sender_celltypes %>% unique() %>% lapply(get_expressed_genes, alldata, 0.10) # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender <- list_expressed_genes_sender %>% unlist() %>% unique()
```

### Define gene set of interest

```{r}
seurat_obj_receiver <- subset(alldata, idents = receiver)
seurat_obj_receiver <- SetIdent(seurat_obj_receiver, value = seurat_obj_receiver[["type"]])

condition_oi <- "Mutant"
condition_reference <- "Wildtype" 
  
DE_table_receiver <- FindMarkers(object = seurat_obj_receiver, ident.1 = condition_oi, ident.2 = condition_reference, min.pct = 0.10) %>% rownames_to_column("gene")

geneset_oi = DE_table_receiver %>% filter(p_val_adj <= 0.05 & abs(avg_log2FC) >= 0.25) %>% pull(gene)
geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```

### Define set of potential Ligands

```{r}
ligands <- lr_network %>% pull(from) %>% unique()
receptors <- lr_network %>% pull(to) %>% unique()

expressed_ligands <- intersect(ligands,expressed_genes_sender)
expressed_receptors <- intersect(receptors,expressed_genes_receiver)

potential_ligands <- lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()
```

### Perform NicheNet ligand activity analysis
Rank the potential ligands based on the presence of their target genes in the gene set of interest (compared to the background set of genes)

```{r}
ligand_activities <- predict_ligand_activities(geneset = geneset_oi, background_expressed_genes = background_expressed_genes, ligand_target_matrix = ligand_target_matrix, potential_ligands = potential_ligands)

ligand_activities <- ligand_activities %>% arrange(-pearson) %>% mutate(rank = rank(desc(pearson)))
ligand_activities

#The number of top-ranked ligands that are further used to predict active target genes and construct an active ligand-receptor network is here 20.

best_upstream_ligands = ligand_activities %>% top_n(40, pearson) %>% arrange(-pearson) %>% pull(test_ligand) %>% unique()

#These ligands are expressed by one or more of the input sender cells. To see which cell population expresses which of these top-ranked ligands, you can run the following:
pdf("/data/user/tchowton/220804_YoderSC/Figures/ift88-nichenet/220826_best-upstream-ligands.pdf", width = 15, height = 10)
DotPlot(alldata, features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis()
dev.off()
```

### Infer receptors and top-predicted target genes of ligands that are top-ranked in the ligand activity analysis
Active target gene inference

```{r}
active_ligand_target_links_df <- best_upstream_ligands %>% lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>% bind_rows() %>% drop_na()

active_ligand_target_links <- prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.33)

order_ligands <- intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
order_targets <- active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()
rownames(active_ligand_target_links) <- rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
colnames(active_ligand_target_links) <- colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23

vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()

p_ligand_target_network <- vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple",legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential")  + theme(axis.text.x = element_text(face = "italic")) + scale_fill_gradient2(low = "whitesmoke",  high = "purple", breaks = c(0,0.0045,0.0090))

pdf("/data/user/tchowton/220804_YoderSC/Figures/ift88-nichenet/220826_ligand-target-heatmap.pdf", width = 15, height = 10)
p_ligand_target_network
dev.off()
```

Receptors of top-ranked ligands

```{r}
lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()

lr_network_top_df_large = weighted_networks_lr %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors)

lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
lr_network_top_matrix = lr_network_top_df %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)

dist_receptors = dist(lr_network_top_matrix, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]
    
dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix))
order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix))

vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
rownames(vis_ligand_receptor_network) = order_receptors %>% make.names()
colnames(vis_ligand_receptor_network) = order_ligands_receptor %>% make.names()

p_ligand_receptor_network = vis_ligand_receptor_network %>% t() %>% make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential")

pdf("/data/user/tchowton/220804_YoderSC/Figures/ift88-nichenet/220826_ligand-receptor.pdf", width = 15, height = 10)
p_ligand_receptor_network
dev.off()
```

Receptors of top-ranked ligands, but after considering only bona fide ligand-receptor interactions documented in literature and publicly available databases

```{r}
lr_network_strict = lr_network %>% filter(database != "ppi_prediction_go" & database != "ppi_prediction")
ligands_bona_fide = lr_network_strict %>% pull(from) %>% unique()
receptors_bona_fide = lr_network_strict %>% pull(to) %>% unique()

lr_network_top_df_large_strict = lr_network_top_df_large %>% distinct(from,to) %>% inner_join(lr_network_strict, by = c("from","to")) %>% distinct(from,to)
lr_network_top_df_large_strict = lr_network_top_df_large_strict %>% inner_join(lr_network_top_df_large, by = c("from","to"))

lr_network_top_df_strict = lr_network_top_df_large_strict %>% spread("from","weight",fill = 0)
lr_network_top_matrix_strict = lr_network_top_df_strict %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df_strict$to)

dist_receptors = dist(lr_network_top_matrix_strict, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]

dist_ligands = dist(lr_network_top_matrix_strict %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix_strict))
order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix_strict))

vis_ligand_receptor_network_strict = lr_network_top_matrix_strict[order_receptors, order_ligands_receptor]
rownames(vis_ligand_receptor_network_strict) = order_receptors %>% make.names()
colnames(vis_ligand_receptor_network_strict) = order_ligands_receptor %>% make.names()

p_ligand_receptor_network_strict = vis_ligand_receptor_network_strict %>% t() %>% make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential\n(bona fide)")

pdf("/data/user/tchowton/220804_YoderSC/Figures/ift88-nichenet/220826_ligand-receptor-strict.pdf", width = 5, height = 5)
p_ligand_receptor_network_strict
dev.off()
```


