---
title: "220822_YoderSC_05_dge"
author: "T.C. Howton"
date: '2022-08-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
    library(Seurat)
    library(dplyr)
    library(cowplot)
    library(ggplot2)
    library(pheatmap)
    library(enrichR)
    library(rafalib)
})
```


### Load data

```{r Load-Data}
# Loading in the filtered seurat object from the clustering script
alldata <- readRDS("/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220829_YoderSC_04-pc2-clustering.rds")

dim(alldata) #20405 20625 pc2 - 19893 10688
print(names(alldata@reductions)) # "pca"          "umap"         "harmony"      "umap_harmony"

# Set the identity as leiden with resolution 5
alldata@meta.data %>% head()
sel.clust = "RNA_snn_res.1.5"

alldata <- SetIdent(alldata, value = sel.clust)
table(alldata@active.ident)


# plot this clustering
pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-clustering/220829_res-1_5.pdf')
DimPlot(alldata, label = T) + NoLegend()
DimPlot(alldata, group.by = "orig.ident") + NoAxes()
DimPlot(alldata, split.by = "type") + NoAxes()
DimPlot(alldata, group.by = "type") + NoAxes()
dev.off()

```

## Cell marker genes

Let us first compute a ranking for the highly differential genes in each cluster. There are many different tests and parameters to be chosen that can be used to refine your results. When looking for marker genes, we want genes that are positivelly expressed in a cell type and possibly not expressed in the others.

### Compute differential expression

```{r Compute-Diff-Exp}
markers_genes <- FindAllMarkers(alldata, log2FC.threshold = 0.2, test.use = "wilcox",
    min.pct = 0.1, min.diff.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 50,
    assay = "RNA")

# We can now select the top 25 up regulated genes for plotting.
markers_genes %>%
    group_by(cluster) %>%
    top_n(-25, p_val_adj) -> top25
top25

# We can now select the top 25 up regulated genes for plotting.

mypar(2, 5, mar = c(4, 6, 3, 1))
for (i in unique(top25$cluster)) {
    barplot(sort(setNames(top25$avg_log2FC, top25$gene)[top25$cluster == i], F),
        horiz = T, las = 1, main = paste0(i, " vs. rest"), border = "white", yaxs = "i")
    abline(v = c(0, 0.25), lty = c(1, 2))
}

# We can visualize them as a heatmap. Here we are selecting the top 5.

markers_genes %>%
    group_by(cluster) %>%
    top_n(-5, p_val_adj) -> top5

# create a scale.data slot for the selected genes
alldata <- ScaleData(alldata, features = as.character(unique(top5$gene)), assay = "RNA")
pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-clustering/220829_dge-heatmap.pdf')
DoHeatmap(alldata, features = as.character(unique(top5$gene)), group.by = sel.clust,
    assay = "RNA")
dev.off()
```

### Save Data
```{r SaveData-SessionInfo}
saveRDS(alldata, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220829_YoderSC_05-pc2-dge.rds")

sessionInfo()
```





























