---
title: "220815_YoderSC_01_qc"
output: html_document
date: '2022-08-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

### Load Packages

```{r Load-Packages, echo=FALSE}
library(Seurat)
```


### Load data

```{r Load-Data}
#KZ1cre-1
ift88_wt_1 <- Read10X(data.dir = "/data/project/lasseigne_lab/DATASET_dir/Yoder_SC_IFT88_PC2/Raw-Data/KZ1cre-1/filtered_feature_bc_matrix/")

#KZ4cre-2
ift88_wt_2 <- Read10X(data.dir = "/data/project/lasseigne_lab/DATASET_dir/Yoder_SC_IFT88_PC2/Raw-Data/KZ4cre-2/filtered_feature_bc_matrix/")

#KZ2Cre+1
ift88_ko_1 <- Read10X(data.dir = "/data/project/lasseigne_lab/DATASET_dir/Yoder_SC_IFT88_PC2/Raw-Data/KZ2Cre+1/filtered_feature_bc_matrix/")

#KZ3Cre+1
ift88_ko_2 <- Read10X(data.dir = "/data/project/lasseigne_lab/DATASET_dir/Yoder_SC_IFT88_PC2/Raw-Data/KZ3Cre+2/filtered_feature_bc_matrix/")
```

### Create Seurat objects
Creating the Seurat objects from the loaded matixes

```{r Create-Seurat-Objects}
wt_1_obj <- CreateSeuratObject(counts = ift88_wt_1, project = 'wt-1')
wt_2_obj <- CreateSeuratObject(counts = ift88_wt_2, project = 'wt-2')
ko_1_obj <- CreateSeuratObject(counts = ift88_ko_1, project = 'ko-1')
ko_2_obj <- CreateSeuratObject(counts = ift88_ko_2, project = 'ko-2')

# add metadata
wt_1_obj$type = "Wildtype"
wt_2_obj$type = "Wildtype"
ko_1_obj$type = "Mutant"
ko_2_obj$type = "Mutant"

# Merge datasets into one single seurat object
alldata <- merge(wt_1_obj, c(wt_2_obj, ko_1_obj, ko_2_obj), add.cell.ids = c("wt_1", "wt_2", "ko_1", "ko_2"))

# remove all objects that will not be used.
rm(ift88_wt_1, ift88_wt_2, ift88_ko_1, ift88_ko_2, wt_1_obj, wt_2_obj, ko_1_obj, ko_2_obj)

# run garbage collect to free up memory
gc()

head(alldata@meta.data, 10)
```

### Calculating QC

```{r Calc-QC}
# Calculate the mitochondrial read percentage
alldata <- PercentageFeatureSet(alldata, "^mt-", col.name = "percent_mito")

# Calculate the ribosamal read percentage
alldata <- PercentageFeatureSet(alldata, "^Rp[sl]", col.name = "percent_ribo")

```

### Plot Initial QC
```{r Plot-Initial-QC}
# Create a list of QC features to plot
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")

# Plot and save the initial QC figure
pdf('/data/user/tchowton/220804_YoderSC/Figures/220815_qc-initial.pdf')
VlnPlot(alldata, group.by = "orig.ident", features = feats, pt.size = 0, ncol = 4) +
    NoLegend()
FeatureScatter(alldata, "nCount_RNA", "nFeature_RNA", group.by = "orig.ident", pt.size = 0.5)
dev.off()
```

### Filtering 

```{r Filtering}
# Filter by cell number and feature number
selected_c <- WhichCells(alldata, expression = nFeature_RNA > 200)
selected_f <- rownames(alldata)[Matrix::rowSums(alldata) > 3]

data.filt <- subset(alldata, features = selected_f, cells = selected_c)
dim(data.filt) # 20522 35354
dim(alldata) # 31053 37576
```

```{r Plotting-total-top-genes}
par(mar = c(4, 8, 2, 1))
C <- data.filt@assays$RNA@counts
C <- Matrix::t(Matrix::t(C)/Matrix::colSums(C)) * 100
most_expressed <- order(apply(C, 1, median), decreasing = T)[20:1]
pdf('/data/user/tchowton/220804_YoderSC/Figures/220815_total-top-genes-filtered.pdf')
boxplot(as.matrix(Matrix::t(C[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)
dev.off()
```

```{r}
selected_mito <- WhichCells(data.filt, expression = percent_mito < 20)
selected_ribo <- WhichCells(data.filt, expression = percent_ribo > 5)

# and subset the object to only keep those cells
data.filt <- subset(data.filt, cells = selected_mito)
data.filt <- subset(data.filt, cells = selected_ribo)

dim(data.filt)

table(data.filt$orig.ident)
```

### Filtering technical genes
As the level of expression of mitochondrial and MALAT1 genes are judged as mainly technical, it can be wise to remove them from the dataset bofore any further analysis.

```{r}
dim(data.filt)

# Filter MALAT1
data.filt <- data.filt[!grepl("Malat1", rownames(data.filt)), ]

# Filter Mitocondrial
data.filt <- data.filt[!grepl("^mt-", rownames(data.filt)), ]

# Filter Ribossomal gene (optional if that is a problem on your data) data.filt
data.filt <- data.filt[ ! grepl('^Rp[sl]', rownames(data.filt)), ]

dim(data.filt)
```

### Cell Cycle Scoring

```{r}
# Before running CellCycleScoring the data need to be normalized and logtransformed.
data.filt = NormalizeData(data.filt)

#reading in cell cycle gene list (downloded from https://github.com/hbc/tinyatlas/blob/master/cell_cycle/Mus_musculus.csv)
cell_cycle_gene <- read.delim("/data/user/tchowton/220804_YoderSC/Data/mouse-cell-cycle-genes.txt", header = FALSE)

data.filt <- CellCycleScoring(object = data.filt, g2m.features = cell_cycle_gene[1:54,],
    s.features = cell_cycle_gene[55:97,])

head(data.filt@meta.data)

#plotting cell cycle scores per sample
VlnPlot(data.filt, 
        features = c("S.Score", "G2M.Score"), 
        group.by = "orig.ident", 
        ncol = 4, 
        pt.size = 0)
```

### Plotting Filtered QC
```{r Plot-filtered-QC}
# Create a list of QC features to plot
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")

# Plot and save the initial QC figure
pdf('/data/user/tchowton/220804_YoderSC/Figures/220816_qc-filtered.pdf')
VlnPlot(data.filt, 
        group.by = "orig.ident", 
        features = feats, 
        pt.size = 0, 
        ncol = 4) +
    NoLegend()
FeatureScatter(alldata, "nCount_RNA", "nFeature_RNA", group.by = "orig.ident", pt.size = 0.5)
VlnPlot(data.filt, 
        features = c("S.Score", "G2M.Score"), 
        group.by = "orig.ident", 
        ncol = 4, 
        pt.size = 0)
dev.off()
```

### Save Data
```{r SaveData-SessionInfo}
saveRDS(data.filt, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220816_YoderSC_01-ift88-qc.rds")

sessionInfo()
```
























