---
title: "220815_YoderSC_01_qc"
output: html_document
date: '2022-08-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
This script will perform quality control for the ift88 mutant samples from Dr. Yoder's lab. 

## Analysis

### Load Packages

```{r Load-Packages, echo=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
})
```

### Create Seurat objects
Creating the Seurat objects from the loaded matixes.

```{r Create-Seurat-Objects}
# Function to read in data and create Seurat objects. Provide the path to the directory containing the cellranger output directories for each sample.
create_seurat_objects <- function(path){
  
  # Generate the raw input list that contains the sample names
  input_list <- list.dirs(path = path, full.names = FALSE, recursive = FALSE)

  #For loop to read in the data from the CellRanger output and create a cooresponding Seurat object. 
  for (sample in input_list) {
    # Reads in the data. the sample variable here is to distinguish between the data directories from the STARsolo output
    seurat_data <- Read10X(data.dir = paste0(path, sample, "/filtered_feature_bc_matrix/"))
    
    # Creates the Seurat object for the sample. Sets the project to the sample name.
    seurat_obj <- CreateSeuratObject(counts = seurat_data,
                                      min.features = 100,
                                      project = sample)
    
    #assigns the seurat object to the sample name.
    assign(sample, seurat_obj, envir = .GlobalEnv)
  }
}

create_seurat_objects(path = "/data/user/tchowton/220804_YoderSC/Data/RawData/")
create_seurat_objects(path = "/data/user/tchowton/PkdMtOxphos/Data/cellRangerOutput/220624_cellRangerOutputsPC2/")


# Add 'type' metadata ift88
ift88_wt_1$type = "Wildtype"
ift88_wt_2$type = "Wildtype"
ift88_ko_1$type = "Mutant"
ift88_ko_2$type = "Mutant"

# Add 'type' metadata ift88
pc2_wt_1$type = "Wildtype"
pc2_wt_2$type = "Wildtype"
pc2_ko_1$type = "Mutant"
pc2_ko_2$type = "Mutant"

# Merge datasets into one single seurat object ift88
alldata_ift88 <- merge(ift88_wt_1,
                 c(ift88_wt_2, ift88_ko_1, ift88_ko_2),
                 add.cell.ids = c("ift88_wt_1", "ift88_wt_2", "ift88_ko_1", "ift88_ko_2"))

# Merge datasets into one single seurat object pc2
alldata_pc2 <- merge(pc2_wt_1,
                 c(pc2_wt_2, pc2_ko_1, pc2_ko_2),
                 add.cell.ids = c("pc2_wt_1", "pc2_wt_2", "pc2_ko_1", "pc2_ko_2"))

# remove all objects that will not be used.
rm(ift88_wt_1, ift88_wt_2, ift88_ko_1, ift88_ko_2, pc2_wt_1, pc2_wt_2, pc2_ko_1, pc2_ko_2)

# run garbage collect to free up memory
gc()

head(alldata@meta.data, 10)
```

### Calculating QC

```{r Calc-QC}
# Calculate the mitochondrial read percentage
alldata_ift88 <- PercentageFeatureSet(alldata_ift88, "^mt-", col.name = "percent_mito")

# Calculate the ribosamal read percentage
alldata_ift88 <- PercentageFeatureSet(alldata_ift88, "^Rp[sl]", col.name = "percent_ribo")

# Calculate the mitochondrial read percentage
alldata_pc2 <- PercentageFeatureSet(alldata_pc2, "^mt-", col.name = "percent_mito")

# Calculate the ribosamal read percentage
alldata_pc2 <- PercentageFeatureSet(alldata_pc2, "^Rp[sl]", col.name = "percent_ribo")

```

### Plot Initial QC
```{r Plot-Initial-QC}
# Create a list of QC features to plot
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")

# Plot and save the initial QC figure
pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-preprocessing/220829_qc-initial.pdf')
VlnPlot(alldata, group.by = "orig.ident", features = feats, pt.size = 0, ncol = 4) +
    NoLegend()
FeatureScatter(alldata, "nCount_RNA", "nFeature_RNA", group.by = "orig.ident", pt.size = 0.5)
dev.off()
```

### Filtering 

```{r Filtering}
# Filter by cell number and feature number
selected_c <- WhichCells(alldata_ift88, expression = nFeature_RNA > 200)
selected_f <- rownames(alldata_ift88)[Matrix::rowSums(alldata_ift88) > 3]

data.filt_ift88 <- subset(alldata_ift88, features = selected_f, cells = selected_c)
dim(data.filt_ift88) # 20522 35354
dim(alldata_ift88) # 31053 37576

# Filter by cell number and feature number
selected_c <- WhichCells(alldata_pc2, expression = nFeature_RNA > 200)
selected_f <- rownames(alldata_pc2)[Matrix::rowSums(alldata_pc2) > 3]

data.filt_pc2 <- subset(alldata_pc2, features = selected_f, cells = selected_c)
dim(data.filt_pc2) # 20522 35354
dim(alldata_pc2) # 31053 37576

# pc2
# [1] 20005 18799
# [1] 32285 19907
```

```{r Plotting-total-top-genes}
par(mar = c(4, 8, 2, 1))
C <- data.filt@assays$RNA@counts
C <- Matrix::t(Matrix::t(C)/Matrix::colSums(C)) * 100
most_expressed <- order(apply(C, 1, median), decreasing = T)[20:1]
pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-preprocessing/220829_total-top-genes-filtered.pdf')
boxplot(as.matrix(Matrix::t(C[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)
dev.off()
```

```{r}
selected_mito <- WhichCells(data.filt_ift88, expression = percent_mito < 20)
selected_ribo <- WhichCells(data.filt_ift88, expression = percent_ribo > 5)

# and subset the object to only keep those cells
data.filt_ift88 <- subset(data.filt_ift88, cells = selected_mito)
data.filt_ift88 <- subset(data.filt_ift88, cells = selected_ribo)

dim(data.filt_ift88)
#[1] 20521 20625

table(data.filt_ift88$orig.ident)
# ift88_ko_1 ift88_ko_2 ift88_wt_1 ift88_wt_2 
#       5047       5444       5058       5076

selected_mito <- WhichCells(data.filt_pc2, expression = percent_mito < 20)
selected_ribo <- WhichCells(data.filt_pc2, expression = percent_ribo > 5)

# and subset the object to only keep those cells
data.filt_pc2 <- subset(data.filt_pc2, cells = selected_mito)
data.filt_pc2 <- subset(data.filt_pc2, cells = selected_ribo)

dim(data.filt_pc2)

table(data.filt_pc2$orig.ident)

# pc2
# [1] 20005 10688
# 
# pc2_ko_1 pc2_ko_2 pc2_wt_1 pc2_wt_2 
#     2789     2817     2438     2644 
```

### Filtering technical genes
As the level of expression of mitochondrial and MALAT1 genes are judged as mainly technical, it can be wise to remove them from the dataset bofore any further analysis.

```{r}
dim(data.filt)

# Filter MALAT1
data.filt_ift88 <- data.filt_ift88[!grepl("Malat1", rownames(data.filt_ift88)), ]

# Filter Mitocondrial
data.filt_ift88 <- data.filt_ift88[!grepl("^mt-", rownames(data.filt_ift88)), ]

# Filter Ribossomal gene (optional if that is a problem on your data) data.filt
data.filt_ift88 <- data.filt_ift88[ ! grepl('^Rp[sl]', rownames(data.filt_ift88)), ]

dim(data.filt_ift88)
#20404 20625


# Filter MALAT1
data.filt_pc2 <- data.filt_pc2[!grepl("Malat1", rownames(data.filt_pc2)), ]

# Filter Mitocondrial
data.filt_pc2 <- data.filt_pc2[!grepl("^mt-", rownames(data.filt_pc2)), ]

# Filter Ribossomal gene (optional if that is a problem on your data) data.filt
data.filt_pc2 <- data.filt_pc2[ ! grepl('^Rp[sl]', rownames(data.filt_pc2)), ]

dim(data.filt_pc2)
#19893 10688

# > mean(data.filt_ift88@meta.data$nCount_RNA)
# [1] 4247.003
# > mean(data.filt_ift88@meta.data$nFeature_RNA)
# [1] 1874.096
# > mean(data.filt_pc2@meta.data$nCount_RNA)
# [1] 4696.562
# > mean(data.filt_pc2@meta.data$nFeature_RNA)
# [1] 2008.494

metadata_ift88 <- data.filt_ift88@meta.data
metadata_pc2 <- data.filt_pc2@meta.data

metadata_ift88$cells <- rownames(metadata_ift88)

metadata_ift88$sample <- NA
metadata_ift88$sample[which(str_detect(metadata_ift88$cells, "^ift88_wt*"))] <- "ctrl"
metadata_ift88$sample[which(str_detect(metadata_ift88$cells, "ko_"))] <- "mut"

metadata_ift88 %>%
  ggplot(aes(x=sample, fill=sample)) + 
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("NCells")


```

### Cell Cycle Scoring

```{r}
# Before running CellCycleScoring the data need to be normalized and logtransformed.
data.filt = NormalizeData(data.filt)

#reading in cell cycle gene list (downloded from https://github.com/hbc/tinyatlas/blob/master/cell_cycle/Mus_musculus.csv)
cell_cycle_gene <- read.delim("/data/user/tchowton/220804_YoderSC/Data/mouse-cell-cycle-genes.txt", header = FALSE)

data.filt <- CellCycleScoring(object = data.filt, g2m.features = cell_cycle_gene[1:54,],
    s.features = cell_cycle_gene[55:97,])

head(data.filt@meta.data)

#plotting cell cycle scores per sample
VlnPlot(data.filt, 
        features = c("S.Score", "G2M.Score"), 
        group.by = "orig.ident", 
        ncol = 4, 
        pt.size = 0)
```

### Plotting Filtered QC
```{r Plot-filtered-QC}
# Create a list of QC features to plot
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")

# Plot and save the initial QC figure
pdf('/data/user/tchowton/220804_YoderSC/Figures/pc2-preprocessing/220829_qc-filtered.pdf')
VlnPlot(data.filt, 
        group.by = "orig.ident", 
        features = feats, 
        pt.size = 0, 
        ncol = 4) +
    NoLegend()
FeatureScatter(alldata, "nCount_RNA", "nFeature_RNA", group.by = "orig.ident", pt.size = 0.5)
VlnPlot(data.filt, 
        features = c("S.Score", "G2M.Score"), 
        group.by = "orig.ident", 
        ncol = 4, 
        pt.size = 0)
dev.off()
```

### Save Data
```{r SaveData-SessionInfo}
saveRDS(data.filt, "/data/user/tchowton/220804_YoderSC/Data/ProcessedData/220829_YoderSC_01-PC2-qc.rds")

sessionInfo()
```
























