---
title: "220908_YoderSC_08_liana"
author: "T.C. Howton"
date: '2022-09-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Overview
This project will use Liana and NichNet to explore cell-cell communication in PKD samples form Dr. Yoder's lab. This project will reproduce Kasi's work with updated NicheNet and adding Liana.

## Script Purpose
LIANA's OmniPath resource was largely compiled from CCC databases with a human focus, as such LIANA works with human gene symbols by default. Nevertheless, one could run LIANA with any other organism, provided a dictionary is available to map human genesymbols to those of species X. This could be achieved using the HomoloGene database via OmnipathR with LIANA's generate_homologs function.


### Load Packages

```{r Load-packages}
suppressPackageStartupMessages({
  library(tidyverse)
  library(liana)
  library(nichenetr)
  library(Seurat)
  library(ggrepel)
  library(cowplot)
  library(OmnipathR)
  library(magrittr)
})
```


```{r }
liana_path <- system.file(package = "liana")
testdata <-
    readRDS(file.path(liana_path, "testdata", "input", "testdata.rds"))

# Check the organisms which are currently available in omnipathR.
show_homologene()
```

